---
output: html_document
editor_options: 
  chunk_output_type: inline
---
Leon County Tornado 4/23/2020

https://nwschat.weather.gov/p.php?pid=202004281348-KTAE-NOUS42-PNSTAE

START DATE:             04/23/2020
START TIME:             06:22 PM EDT
START LOCATION:         2 SW LAKE JACKSON / LEON COUNTY / FL
START LAT/LON:          30.5181 / -84.3538

END DATE:               04/23/2020
END TIME:               06:38 PM EDT
END LOCATION:           1 WSW PISGAH CHURCH / LEON COUNTY / FL
END LAT/LON:            30.5495 / -84.1755

SURVEY SUMMARY:
A TORNADO TOUCHED DOWN JUST SOUTH OF NORTH MONROE STREET AND
MOVED EAST INTO A NEIGHBORHOOD WHERE THERE WAS WIDESPREAD LIMB
AND TREE DAMAGE. IT LIKELY THEN CROSSED OVER THE LAKE AND
IMPACTED ELINOR KLAPP PHIPPS PARK WHERE SNAPPED TREE TRUNKS WERE
OBSERVED. IT PRODUCED SOME OF THE MOST WIDESPREAD TREE DAMAGE,
INCLUDING SNAPPED TRUCKS, NEAR OX BOTTOM ROAD BEFORE CROSSING
THOMASVILLE ROAD.  IT MOVED NORTHEAST THROUGH KILLEARN ESTATES
AND KILLEARN ACRES WITH ADDITIONAL WIDESPREAD TREE AND LIMB
DAMAGE AND A FEW SPOTS OF SNAPPED AND UPROOTED TREES. THE LAST
SIGNIFICANT DAMAGE WAS ADDITIONAL SNAPPED TREES ON BRADFORDVILLE
ROAD. ADDITIONAL TREE DAMAGE WAS SPORADIC TO THE EAST OF HERE AS
IS LIKELY OUTFLOW AS THE TORNADO LIFTED. THE HIGHEST DAMAGE WAS
RATED 95 MPH DUE TO SNAPPED TREES BUT THE MAJORITY OF THE DAMAGE
WAS EF0 TYPE TREE DAMAGE. 

## Estimate forward speed

Packages
```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

### Single tornado

Create a simple feature data frame with start and end points as separate POINT geometries.
```{r}
Tor.df <- data.frame(id = "TLH",
                     time = c("2020-04-23 18:22:00 EDT", "2020-04-23 18:38:00"),
                     lat = c(30.5181, 30.5495), 
                     lon = c(-84.3538, -84.1755)) %>%
  mutate(time = as_datetime(time, 
                            tz = "America/New_York"))
```

```{r}
Tor.sf <- st_as_sf(Tor.df,
                   coords = c("lon", "lat"),
                   crs = 4326) %>%
  st_transform(crs = 32616)
ds <- as.numeric(st_distance(Tor.sf)[1, 2])
dt <- as.numeric(difftime(Tor.sf$time[2], Tor.sf$time[1], units = "secs"))
ds/dt  # speed in m/s
```

### Multiple tornadoes 

Read storm data. ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/
```{r}
SD2013.df <- read.csv(file = "stormdata_2013.csv")
SD2010.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/stormdata_2010.csv")
SD2011.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/stormdata_2011.csv")
SD2012.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/stormdata_2012.csv")

SD2009.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/Stormdata_2009.csv")  # Capitalized S before 2010
SD2008.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/Stormdata_2008.csv")
SD2007.df <- read.csv(file = "ftp://ftp.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/legacy/Stormdata_2007.csv") %>% 
  filter(MONTH_NAME != "January") # EF scale implementated after January 2007

SD.df <- bind_rows(SD2010.df, SD2011.df, SD2012.df, SD2013.df)

SD2.df <- SD.df %>%
  filter(EVENT_TYPE == "Tornado") %>%
  select(TOR_F_SCALE, TOR_LENGTH, TOR_WIDTH, BEGIN_DATE_TIME, END_DATE_TIME) %>%
  mutate(PathLength = TOR_LENGTH * 1609.344,
         Begin = mdy_hms(BEGIN_DATE_TIME, tz = "America/New_York"), # hms starting with stormdata_2010
         End = mdy_hms(END_DATE_TIME, tz = "America/New_York"),
          Duration = as.numeric(difftime(End, Begin, units = "secs")),
         ForwardSpeed = PathLength/Duration) %>%
  filter(!is.na(ForwardSpeed) & Duration > 300)

SD1.df <- bind_rows(SD2007.df, SD2008.df, SD2009.df)

SD3.df <- SD1.df %>%
  filter(EVENT_TYPE == "Tornado") %>%
  select(TOR_F_SCALE, TOR_LENGTH, TOR_WIDTH, BEGIN_DATE_TIME, END_DATE_TIME) %>%
  mutate(PathLength = TOR_LENGTH * 1609.344,
         Begin = mdy_hm(BEGIN_DATE_TIME, tz = "America/New_York"), # hm before 2010
         End = mdy_hm(END_DATE_TIME, tz = "America/New_York"),
         Duration = as.numeric(difftime(End, Begin, units = "secs")),
         ForwardSpeed = PathLength/Duration) %>%
  filter(!is.na(ForwardSpeed) & Duration > 300)

SD4.df <- bind_rows(SD3.df, SD2.df)

summary(SD3.df$ForwardSpeed)
dim(SD2.df)
quantile(SD2.df$ForwardSpeed, probs = c(.9, .95, .99, .995))
```

```{r}
SD4.df %>%
  group_by(TOR_F_SCALE) %>%
  summarize(N = n(),
            medianFS = round(median(ForwardSpeed), 1),
            meanFS = round(mean(ForwardSpeed), 1)) %>%
  rename(Damage = TOR_F_SCALE)

library(gt)

SD4.df %>%
  group_by(TOR_F_SCALE) %>%
  summarize(N = n(),
            Median = round(median(ForwardSpeed), 1),
            Mean = round(mean(ForwardSpeed), 1)) %>%
  rename(Damage = TOR_F_SCALE) %>%
  gt() %>%
  tab_header(
    title = "Tornado Forward Speeds (m/s)",
    subtitle = "2007-2013"
  ) %>%
  fmt_markdown(
    columns = vars(Damage)
  ) %>%
  fmt_number(
    columns = vars(N, Median, Mean),
    drop_trailing_zeros = TRUE
  )


SD4.df %>%
  mutate(Month = month(Begin)) %>%
  group_by(Month) %>%
  summarize(nT = n(),
            medianFS = median(ForwardSpeed),
            meanFS = mean(ForwardSpeed))
```
